\large \textbf{Title:} Computational models reveal how cleaner fish adjust decisions in a biological market

### 2022-07-25 manuscript_BE/manuscript_BE.tex
Further changes to the manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage{graphicx}
\usepackage{booktabs}
\linespread{2}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}\doublespacing
\newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

@@

The prediction error (Eq. \ref{eq:delta}) is used to update the value of
each one of the states the agent faces, as well as the preference for
the resident and visitor options. The value update is simple the product
of the prediction error and the parameter for the speed of learning
(\(\Delta V(S_t)=\alpha \delta_t\)). The change in preference between
the resident and visitor is given by
\(\Delta (\theta_v-\theta_r)_t=\alpha\delta_t2(1-\pi_v)\), where
\(\theta_i\) represent the preference for one of two options and the
difference captures the total change relative to one another; \(\pi_v\)
corresponds to the current probability of choosing the visitor. \(p_i\)
is determined by applying the logistic function to the diffence in
preferences between the two mutually exclusive options
(\(\pi_v=\frac{1}{1+e^{-(\theta_v-\theta_r)}}\)). This amounts to a
preference update that is carried in the direction that leads to more
reward being obtained, given the new information. In the long run, the
probability of choosing a visitor over a resident converges in the
model. To which probability the model will converge depends on the
relative abundance of cleaners, visitors and residents; as well as on
the probability of visitors leaving the cleaning station when made to
wait. Further details of the model implementation can be found in
Quiñones et al. (2019).

The model shows that agents need to find a way to incorporate future
consequences of current choices. In the model, this could be achieved
with either of two parameters that could also work together. First,
\(\gamma\) measures how much individuals include future rewards in their
decision updates. If \(\gamma=0\), individuals only use the immediate
reward obtained from a cleaning interaction. As \(\gamma\) increases,
individuals include more the reward obtained from the subsequent
choices. That amounts to estimating and using for decision making the
future expected rewards of an action (chaining). Second, \(\eta\)
measures how much individuals include in their reward the fleeing
behavior of visitors as a negative component (penalty). Both of these
parameters allow individuals to use in their estimates the future
effects of their choices.

@@

The empirical data were collected between 2010 and 2019 always during
the austral winter months June to August from a total of five study reef
sites (Corner Beach-CB, Horseshoe-HS, Mermaid Cove-MC, Northern
Horseshoe-NHS, and The Crest-TC) at Lizard Island
(\(14.6682^\circ S, 145.4604^\circ E\)), Great Barrier Reef, Australia.
The data consist of three sets: fish censuses, field observations of
cleaner-client interactions to quantify the probability of visitors
leaving if made to wait, and the performance of wild-caught cleaner fish
in the ephemeral reward test. In total, we have twelve site/year data
sets for fish censuses and corresponding performance in the lab test.
Thus, some sites were sampled more than once. To estimate the population
density of cleaner fish and their clients at a given site in a given
year, Triki et al. (2019) used a series of ten transects of \(30m\)
each. Observers swam along the transect lines placed on the reef and
first counted the visible large-bodied adult fish (species with total
length \(TL \geq 10cm\)) including cleaner fish on a width of \(5m\),
and then on the return individuals of small-bodied fish species
(\(TL < 10 cm\)) on a width of \(1 m\) (see Triki et al. (2018) for
further details on fish censuses data collection). We then scaled the
counts of cleaner fish, small-bodied, and large-bodied clients fish
densities per \(100 m^2\). From the study by Triki et al. (2019) cleaner
fish and large-bodied client populations densities were highly
correlated, and only the former was hence used in the analyses as
representative of both measures.

The field observation data consisted of video recordings/encodings of
the cleaner-client cleaning interactions. There were videos from eight
cleaners per site/year of a duration of \(30 min\) each. Triki et al.~
extracted information from every event wherein a visitor client was made
to wait in favour of another client (visitor or resident), and noted
whether or not the visitor left or queued for the cleaning service
(Triki et al. 2019, 2020).

The cognitive performance data was from a total of 120 cleaners (10
individuals per 12 site/year) tested in the ephemeral reward task (Triki
et al. 2019, 2020). Authors housed all captured cleaners individually in
glass aquaria (\(62cm \times 27cm \times 37 cm\)) and provided them with
PVC pipes (\(10 cm \times 1 cm\)) as shelters. The task consisted of
exposing the cleaner fish to substitute models of client fish in the
form of two \emph{Plexiglas} plates offering the same amount of food
(one item of mashed prawn). The two plates differed in colour and
pattern (horizontal green stripes or vertical pink stripes) but had
equal size (\(10 cm \times 7 cm\)). Importantly, the two plates played
different roles as either a visitor (ephemeral food source) or resident
(permanent food source). That is, if a cleaner fish inspected the
resident plate first, the experimenter withdrew the visitor plate out of
the aquarium as a consequence. Choosing first the visitor plate,
however, granted access to both plates. The equal size of the plates
forced cleaner fish to learn to give service priority to the visitor
plate based solely on the behavior-cue of the plates rather than
size-cue (Wismer et al. 2019). Triki et al. (2019, 2020) tested the fish
for a maximum of 200 trials with 20 trials a day, 10 trials in the
morning and 10 trials in the afternoon. They randomized and
counterbalanced the plates' spatial location (i.e.~left or right)
between trials. Similarly, they counterbalanced the plates' decoration
(colour and pattern) and the plates' role (visitor or resident) between
the tested fish. In the original studies, once a fish reach a learning
criterion, that is, performing significantly above chance level (\(> 50%
\), \(p-value ≥ 0.05\)), they passed to a reversal version of the task
where the roles of the visitor/resident Plexiglas plates were swapped
(see Triki et al.(2019, 2020)). Here, we used instead a subset of these
data in order to have an idea of cleaner fish preferences for the
visitor plate, even if they do not reach the learning criterion within
200 trials. To do so, we first extracted the trial-by-trial outcomes
from the last two sessions (20 trials) of those who never reached the
learning criterion for visitor plate (\(N = 45\) cleaner fish). For
those who reached the learning criterion at some point during the test
and passed to a reversal phase, we extracted the trial-by-trial outcomes
from the last session (10 trials) before passing to reversal and the
last 10 trials they were exposed to in the test (\(N = 75\) cleaner
fish). We chose a combination of initial and reversal to quantify
preference for the visitor client, given that it matches well the
criteria chosen in previous analysis of the ephemeral reward task
experimental set-up (Triki et al. 2019, 2020). In the supplementary
material (Fig. \ref{fig:rawdata}) we show how our choice maps to the
previously used criteria. For comparison, we also show a quantification
of visitor preference based only on the initial phase of the task.

@@

The aim of the analyses is to fit the key model parameters \(\gamma\)
and \(\eta\), to the empirical data from Triki et al.~ (2019, 2020) to
test whether each or a combination of these effects is a better
explanation for the pattern seen in the data. We used the ecological
variables: cleaners, visitor clients, resident clients abundances and
visitor clients leaving probability,\\
as input to the models. As the response variable, we used cleaners'
preference for visitors over residents in the ephemeral reward task.
Finally, we used the preference for visitors resulting from the model
simulations as the prediction for the response variable. We kept all
other parameter values used for the model simulations constant, see
Table \ref{tab:param}.

@@

Once we calculated the relative abundances, we obtained predictions from
the model for each one of the locations and ran the Markov Chain. We
started the chain with random values for the parameters of interest,
then ran the computational model once for every reef site using as input
the ecological explanatory variables. The model outputs the probability
of choosing the visitor for each location \(p_i\), where \(i\) is the
index for the 12 locations. Assuming a binomial distribution, the
probability that each of the cleaner fish 20 choices in the ephemeral
reward task was generated by the model is given by
\(\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}\); where \(k_{ij}\) is
the number of times that cleaner fish \(j\) in reef site \(i\) chose the
visitor over the resident; and \(n=20\) (due to our choice of using 20
choices per cleaner fish). By taking the natural logarithm and summing
over all the cleaner fish and reef sites, we obtained the log-likelihood
of the data given the model and parameters. We then proposed a new set
of parameters drawn from a uniform distribution centred around the old
parameter set. The amplitude of the uniform distribution used for each
parameter can be found in Table \ref{tab:param}. Subsequently, we ran
the model and calculated the likelihood with the new parameter set. We
then used the ratio of the two likelihoods to choose which parameter set
to keep. New parameter sets with a higher likelihood than the old set
replaced old ones, and those with a lower likelihood replaced current
ones with a probability equal to the log-likelihood ratio. Given that we
only used the likelihoods in the decision, we used an uninformative
prior. Once we decided whether the new parameter set would replace the
old one, we ran the model again to sample the likelihood distribution of
the parameter set. We then started the cycle again by proposing a new
set of parameters and repeated the process for \(1e^5\) steps. We ran 5
independent chains, discarded the first \(1000\) samples of each chain
as burn in, and after that, we kept 1 in every 100 samples to avoid
autocorrelation. The collection of parameters sets kept in all chains
approximates the posterior distribution. We coded the model as well as
the fitting algorithm in \emph{c++}, the diagnostics and visualization
in R (R Core Team 2021). All codes are accessible at \textless DOI:
10.5281/zenodo.6338073\textgreater{}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_BE_files/figure-latex/rawdata-1} \caption{Relation between the response variable used in this study and the criteria used in previous studies to assess performance in the ephemeral reward task. In the x axis, we classified the performance of cleaner fish according to whether they developed a preference for the visitor in the initial round, in the initial and reversal, or none of them. In the y axis, we add the choices of two experimental sessions: panel on the left uses one session from the initial round and one from the reversal round when possible (as described in the main text); panel on the right uses two sessions from the initial round for all fish.}\label{fig:rawdata}
\end{figure}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage{graphicx}
\usepackage{booktabs}
\linespread{2}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}\doublespacing
\newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

@@

The prediction error (Eq. \ref{eq:delta}) is used to update the value of
each one of the states the agent faces, as well as the preference for
the resident and visitor options. The value update is simple the product
of the prediction error and the parameter for the speed of learning
(\(\Delta V(S_t)=\alpha \delta_t\)). The change in preference between
the resident and visitor is given by
\(\Delta (\theta_v-\theta_r)_t=\alpha\delta_t2(1-\pi_v)\), where
\(\theta_i\) represent the preference for one of two options and the
difference captures the total change relative to one another; \(\pi_v\)
corresponds to the current probability of choosing the visitor. \(p_i\)
is determined by applying the logistic function to the difference in
preferences between the two mutually exclusive options
(\(\pi_v=\frac{1}{1+e^{-(\theta_v-\theta_r)}}\)). This amounts to a
preference update that is carried in the direction that leads to more
reward being obtained, given the new information. In the long run, the
probability of choosing a visitor over a resident converges in the
model. To which probability the model will converge depends on the
relative abundance of cleaners, visitors and residents; as well as on
the probability of visitors leaving the cleaning station when made to
wait. Further details of the model implementation can be found in
Quiñones et al. (2019).

The model shows that agents need to find a way to incorporate future
consequences of current choices. In the model, this could be achieved
with either of two parameters that could also work together. First,
\(\gamma\) measures how much individuals include future rewards in their
decision updates. If \(\gamma=0\), individuals only use the immediate
reward obtained from a cleaning interaction. As \(\gamma\) increases,
individuals include more of the reward obtained from the subsequent
choices. That amounts to estimating and using for decision making the
future expected rewards of an action (chaining). Second, \(\eta\)
measures how much individuals include in their reward the fleeing
behavior of visitors as a negative component (penalty). Both of these
parameters allow individuals to use in their estimates the future
effects of their choices.

@@

The empirical data were collected between 2010 and 2019 always during
the austral winter months June to August from a total of five study reef
sites (Corner Beach-CB, Horseshoe-HS, Mermaid Cove-MC, Northern
Horseshoe-NHS, and The Crest-TC) at Lizard Island
(\(14.6682^\circ S, 145.4604^\circ E\)), Great Barrier Reef, Australia.
The data consist of three sets: fish censuses, field observations of
cleaner-client interactions to quantify the probability of visitors
leaving if made to wait, and the performance of wild-caught cleaner fish
in the ephemeral reward test. In total, we have twelve site/year data
sets for fish censuses and corresponding performance in the lab test.
Thus, some sites were sampled more than once. To estimate the population
density of cleaner fish and their clients at a given site in a given
year, Triki et al. (2019) used a series of ten transects of \(30m\)
each. Observers swam along the transect lines placed on the reef and
first counted the visible large-bodied adult fish (species with total
length \(TL \geq 10cm\)) including cleaner fish on a width of \(5m\),
and then on the return individuals of small-bodied fish species
(\(TL < 10 cm\)) on a width of \(1 m\) (see Triki et al. (2018) for
further details on fish censuses data collection). Total length
estimates were done by the observer. We then scaled the counts of
cleaner fish, small-bodied, and large-bodied clients fish densities per
\(100 m^2\).

The field observation data consisted of video recordings/encodings of
the cleaner-client cleaning interactions. There were videos from eight
cleaners per site/year of a duration of \(30 min\) each. Triki et al.~
extracted information from every event wherein a visitor client was made
to wait in favor of another client (visitor or resident), and noted
whether or not the visitor left or queued for the cleaning service
(Triki et al. 2019, 2020).

The cognitive performance data was from a total of 120 cleaners (10
individuals per 12 site/year) tested in the ephemeral reward task (Triki
et al. 2019, 2020). Authors housed all captured cleaners individually in
glass aquaria (\(62cm \times 27cm \times 37 cm\)) and provided them with
PVC pipes (\(10 cm \times 1 cm\)) as shelters. The task consisted of
exposing the cleaner fish to substitute models of client fish in the
form of two \emph{Plexiglas} plates offering the same amount of food
(one item of mashed prawn). The two plates differed in colour and
pattern (horizontal green stripes or vertical pink stripes) but had
equal size (\(10 cm \times 7 cm\)). Importantly, the two plates played
different roles as either a visitor (ephemeral food source) or resident
(permanent food source). That is, if a cleaner fish inspected the
resident plate first, the experimenter withdrew the visitor plate out of
the aquarium as a consequence. Choosing first the visitor plate,
however, granted access to both plates. The equal size of the plates
forced cleaner fish to decide based solely on the association between
the behaviour and the collor/pattern cue (Wismer et al. 2019). Triki et
al. (2019, 2020) tested the fish for a maximum of 200 trials with 20
trials a day, 10 trials in the morning and 10 trials in the afternoon.
They randomized and counterbalanced the plates' spatial location
(i.e.~left or right) between trials. Similarly, they counterbalanced the
plates' decoration (colour and pattern) and the plates' role (visitor or
resident) between the tested fish. In the original studies, once a fish
reached a learning criterion, that is, performing significantly above
chance level in a binomial test (\(p-value \leq 0.05\), ), they passed
to a reversal version of the task where the roles of the
visitor/resident Plexiglas plates were swapped. The reversal phased
stopped when the fish performed significatly above chance, or the fish
completed the 200 trial together with the initial; see Triki et
al.(2019, 2020). Here, we are interested in explaining the total
frequency of visitor choices using the model, rather than just the
achievement of the criterion. Total frequency of visitor choices
naturally comes out of the model, and allows us to use all the variation
among cleaners, instead of reducing that to a binomial variable. Thus,
we used instead a subset of these data to estimate the final cleaner
fish preferences for the visitor plate, even if they do not reach the
learning criterion within 200 trials. To do so, we first extracted the
trial-by-trial outcomes from the last two sessions (20 trials) of those
who never reached the learning criterion for visitor plate (\(N = 45\)
cleaner fish). For those who reached the learning criterion at some
point during the test and passed to a reversal phase, we extracted the
trial-by-trial outcomes from the last session (10 trials) before passing
to reversal and the last 10 trials they were exposed to in the test
(\(N = 75\) cleaner fish).

We chose a combination of initial and reversal to quantify preference
for the visitor client. However, it could be argued that using only the
initial phase gives a better estimation of the cleaner fish preference
for the visitor. In the supplementary material (Fig. \ref{fig:rawdata})
we show how using initial and reversal (a), and only initial (b) maps to
the previously used criteria. The initial and reversal match better the
criteria chosen in previous analysis of the ephemeral reward task
experimental set-up (Triki et al. 2019, 2020).

@@

The aim of the analyses is to fit the key model parameters \(\gamma\)
and \(\eta\), to the empirical data from Triki et al.~ (2019, 2020) to
test whether each or a combination of these effects is a better
explanation for the pattern seen in the data. We used the ecological
variables: cleaners, visitor clients, resident clients abundances and
visitor clients leaving probability,\\
as input to the models. As the response variable, we used the frequency
with which cleaners chose the visitor option in the ephemeral reward
task. Finally, we used the probability with which agents in the model
simulations choose the visitor in resident-visitor options as the
prediction for the response variable. We kept all other parameter values
used for the model simulations constant, see Table \ref{tab:param}.

@@

Once we calculated the relative abundances, we obtained predictions from
the model for each one of the locations and ran the Markov Chain. We
started the chain with random values for the parameters of interest,
then ran the computational model once for every reef site using as input
the ecological explanatory variables. The model outputs the probability
of choosing the visitor for each location \(p_i\), where \(i\) is the
index for the 12 locations. Assuming a binomial distribution, the
probability that each of the cleaner fish 20 choices in the ephemeral
reward task was generated by the model is given by
\(\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}\); where \(k_{ij}\) is
the number of times that cleaner fish \(j\) in reef site \(i\) chose the
visitor over the resident; and \(n=20\) (due to our choice of using 20
choices per cleaner fish). By taking the natural logarithm and summing
over all the cleaner fish and reef sites, we obtained the log-likelihood
of the data given the model and parameters. We then proposed a new set
of parameters drawn from a uniform distribution centred around the old
parameter set. The amplitude of the uniform distribution used for each
parameter can be found in Table \ref{tab:param}. Subsequently, we ran
the model and calculated the likelihood with the new parameter set. We
then used the ratio of the two likelihoods to choose which parameter set
to keep. New parameter sets with a higher likelihood than the old set
replaced old ones, and those with a lower likelihood replaced current
ones with a probability equal to the log-likelihood ratio. Given that we
only used the likelihoods in the decision, we used an uninformative
prior. Once we decided whether the new parameter set would replace the
old one, we ran the model again to sample the likelihood distribution of
the parameter set. We then started the cycle again by proposing a new
set of parameters and repeated the process for \(1e^5\) steps. We ran 5
independent chains, discarded the first \(1000\) samples of each chain
as burn in, and after that, we kept 1 in every 100 samples to avoid
autocorrelation. The collection of parameters sets kept in all chains
approximates the posterior distribution. We coded the model as well as
the fitting algorithm in \emph{c++}, the diagnostics and visualization
in R (R Core Team 2021). All codes are accessible at
\href{https://zenodo.org/badge/latestdoi/440585701}{\includegraphics{https://zenodo.org/badge/440585701.svg}}.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_BE_files/figure-latex/rawdata-1} \caption{Relation between the response variable used in this study and the criteria used in previous studies to assess performance in the ephemeral reward task. In the x axis, we classified the performance of cleaner fish according to whether they developed a preference for the visitor in the initial round (initial only), in the initial and reversal (both), or none of them (none). In the y axis, we add the choices of two experimental sessions: in panel \textbf{a} we use one session from the initial round and one from the reversal round when possible (as described in the main text); in panel \textbf{b} we use two sessions from the initial round for all fish.}\label{fig:rawdata}
\end{figure}
